Site-Heterogeneous Character Change Models for Morphology

### 2017-09-14 Papers/DraftOutline_SSCCM.tex
adding manuscript updates

@@

	This is a stark contrast to molecular data, in which base pairs or nucleotide residues are generally assumed to have similar properties across an alignment. \par

@@

MrBayes implemented a parameter called the Symmetric Dirichlet Hyperprior, which allowed users to place a prior on state frequencies. 
The probability of observing a change in a character is dependent not only on the probability of change from this character to another, but on the frequency of that character.

@@

This analysis was implemented as a model extension in MrBayes.
In the case of binary data, this was a Beta prior (the Dirichlet is a generalization of the Beta distribution), which operated in a fashion similar to gamma-distributed rate variation: the distribution is discretized into a user-specified number of categories, the median forward and backward transition rate are calculated for each category, and the likelihood of the character is calculated over each category and summed to form the total likelihood. 

@@

In RevBayes, we have implemented a number of useful extensions to the Dirichlet prior of RevBayes, and a constrained version of the CAT model. 

@@

To test the efficacy of our analytical techniques, we use an ant dataset created from datasets from Barden and Grimaldi and one from Keller.
Ants are ecologically crucial organisms as both interacting partners for a variety of plants and animals, and as shapers of the ecosystem via soil cycling and nest building.
As such, they have attracted much work in the world of molecular systematics.
Ants have also have a rich fossil record, with most subfamilies being represented. \par
The monophyly of major subfamilies has been consistently supported by most molecular and morphological studies, with the exception of the Cerapachyinae. 
Molecular systematics had originally shaken up the ant tree of life considerably, breaking up clades considered to be monophyletic based on morphological work.
However, recent family-wide molecular systematic studies provided support for previous morphological work, supporting a clade of subfamilies Ectatomminae and Heteroponerinae, with Myrmicinae as sister to this clade. \par
Excellent morphological matrices on the ingroup of the Formicidae have been available since the early '90s. 
Recent morphological matrices have expanded the sampling of ants from in the large Ponerine family, which was previously undersampled.
Sampling of fossil ants was also expanded to include specimens from the Cretaceous.
These ants were concluded to be stem lineages, and some could not be concretely placed within lineages. \par
In this study, we combine the extant matrices of Keller, and the extinct-extant matrix of Barden and Grimaldi to expand the taxon and character sampling.  
This dataset has interesting properties.
Because there are stem ant lineages represented, there are taxa that have characteristics that are lost after the divergence of the stem lineages.
This extremely one-sided loss structure violates the Mk model assumption of equal transition rates. 
We would expect for these characters to be better modeled by a model that can accommodate asymmetrical transition rates.
Some apomorphies of the ant group are also gained after the divergence of the stem lineages, which also violates the assumption of equal change probabilities.
Because of these model violations, this dataset is an excellent test case for models that relax assumptions of the Mk model. \par
We use Bayes Factor model selection to assess the fit of these relaxed models to the data.
Using this dataset, we strongly support that the use of models that relax key assumptions of the Mk model can greatly improve the fit of the model to the data.
Some stuff about trees and parameters!

@@

	a) Bayes Factor model fitting
	b) TreeSet viz showing how use of the asymetric beta changes our pattern of exploration of treespace

Results

@@

Discussion

@@

>>>>

	For example, in one character, changing from a `0' state to a `1' state could represent a change at a single locus, but at another could represent a much larger change coordinated across many genes.

@@

	This is a stark contrast to molecular data, in which base pairs or amino acid residues are generally assumed to have similar properties across an alignment. \par

@@

MrBayes implemented a parameter called the Symmetric Dirichlet Hyperprior, which allowed users to place a prior or hyperprior on state frequencies. 
The probability of observing a change in a character is dependent not only on the probability of change from this character to another, but on the frequency of the starting character.

@@

In the case of binary data, the prior used was a Beta prior (the Dirichlet is a generalization of the Beta distribution), which operated in a fashion similar to gamma-distributed rate variation: the distribution is discretized into a user-specified number of categories, the median forward and backward transition rate are calculated for each category, and the likelihood of the character is calculated over each category and summed to form the total likelihood. 

@@

In RevBayes, we have implemented a number of useful extensions to the Dirichlet prior of MrBayes, and a constrained version of the CAT model. 

@@

To test the efficacy of our analytical techniques, we use an ant dataset created from datasets from Barden and Grimaldi and one from Keller.
Ants are ecologically crucial organisms as both interacting partners for a variety of plants and animals, and as shapers of the ecosystem via soil cycling and nest building.
As such, they have attracted much work in the world of molecular systematics.
Ants have also have a rich fossil record, with most subfamilies being represented. 
This fossil record has been used for systematic work, as well as a variety of other ecological and evolutionary questions.\par
The monophyly of major subfamilies has been consistently supported by most molecular and morphological studies, with the exception of the Cerapachyinae. 
Molecular systematics had originally shaken up the ant tree of life considerably, breaking up clades considered to be monophyletic based on morphological work.
For example, based on morphological work, six current subfamilies had been previously considered to be a single subfamily, the Ponerinae, until molecular evidence indicated that two of the clades (Ectatomminae and Heteroponerinae) within that subfamily were demonstrated to be more closely related to other ants on the tree.
The Ponerinae was then broken into six distinct subfamilies, one of which is called Ponerinae. 
Recent molecular work has continued to support these six subfamilies as monophyletic, though their relationships to one another remain poorly supported. \par
Excellent morphological matrices on the ingroup of the Formicidae have been available since the early '90s. 
Recent morphological matrices have expanded the sampling of ants from in the large Ponerine family, which was previously undersampled.
Sampling of fossil ants was also expanded to include specimens from the Cretaceous.
These ants were concluded to be stem lineages, and some did not demonstrate any particular taxonomic affinity. 
Even with expanded taxon sampling in the Ponerinae, there is still conflict between the molecular and morphological estimates of phylogenetic relationships. \par
In this study, we combine the extant matrices of Keller, and the extinct-extant matrix of Barden and Grimaldi to expand the taxon and character sampling.  
This dataset has interesting properties.
Because there are stem ant lineages represented, there are taxa that have characteristics that are lost after the divergence of the stem lineages.
This extremely one-sided loss structure violates the Mk model assumption of equal transition rates. 
We would expect for these characters to be better modeled by a model that can accommodate asymmetrical transition rates.
Some apomorphies of the ant group are also gained after the divergence of the stem lineages, which also violates the assumption of equal change probabilities.
Because of these model violations, this dataset is an excellent test case for models that relax assumptions of the Mk model. \par
We use Bayes Factor model selection to assess the fit of these relaxed models to the data.
Using this dataset, we strongly support that the use of models that relax key assumptions of the Mk model can greatly improve the fit of the model to the data.
We also perform simulations to demonstrate that the true number of transition rate asymmetry categories is detectable from the data.

@@

\subsection{Data Matrices}
\subsubsection{Empirical Matrices}
Several large and well-documented ant matrices were used in this study.
The first was that of Keller (2011). 
This matrix is of extant ant groups. 
It was collected for , and has characters and taxa.
Of the total character set, were binary and were multistate.
Because of the scope of Keller's study, the taxon sampling is biased towards and away from . \par
We also used data from Barden and Grimaldi (2015). 
This matrix contained both extant and extinct ants, and was collected in order to place a new taxon of ants from the period.
The matrix contained characters and taxa. \par

In order to achieve maximum coverage, we merged these character matrices.
The taxonomic overlap between the two was taxa, and the character overlap (those characters represented in both matrices without any recoding) was .
The Barden matrix contained characters that were inapplicable to the Keller matrix.
The Keller matrix contained characters that were innapplicable to the Barden matrix.
Inapplicable characters were left in the matrix due to their ability to resolve bipartitions in the the authors' respective groups of interest. \par
Six characters were recoded to bring characters from the original matrices into line with one another.
Most of these changes simply involved changing the terminology used in reference to the character states.
For example, Barden and Grimaldi's character 15 is the same as Keller's character 47, but the two matrices had inverse character codings relative to one another.
The remainder of the changes were changing binary characters to a multistate in cases where one author had more states than the other. \par

	b) Multi-state data matrix
		b1) Model Fit
		b2) Topology
	c) simulated data sets
		c1) Detectability
		c2) Correctness

\section{Results}

@@

\section{Discussion}

@@

\section{Conclusion}

\section{Ackno}
